Sentence-Matching in the Bluegrass Corpus: Methodological Considerations

### 2019-08-16 article.Rmd
small edits

make_diff_matrix <- function(pair, trim) {
  a <- pair %>% 
    filter(speaker_id == "Foreign") %>% 
    select(ends_with("rating")) %>% 
    t() %>% 
    as.matrix()
  b <- pair %>% 
    filter(speaker_id == "Native") %>% 
    select(ends_with("rating")) %>% 
    t() %>% 
    as.matrix()
  mat_statistics <- matrix(0, nrow = ncol(a), ncol = ncol(b))
  mat_pvals <- matrix(0, nrow = ncol(a), ncol = ncol(b))
  for (i in 1:ncol(a)) {
    for (j in 1:ncol(b)) {
      result <- profile_diff(x = a[, i], y = b[, j], trim = trim)
      mat_statistics[i, j] <- result$statistic
      mat_pvals[i, j] <- result$p_value
    }
  }
  rownames(mat_statistics) <- pair %>% 
    filter(speaker_id == "Foreign") %>%
    pull(sentence_id)
  colnames(mat_statistics) <- pair %>% 
    filter(speaker_id == "Native") %>%
    pull(sentence_id)
  rownames(mat_pvals) <- pair %>% 
    filter(speaker_id == "Foreign") %>%
    pull(sentence_id)
  colnames(mat_pvals) <- pair %>% 
    filter(speaker_id == "Native") %>%
    pull(sentence_id)
  list(
    mat_statistics = mat_statistics,
    mat_pvals = mat_pvals
  )
}

@@

  ## get sentence ids for each speaker:
  a_sentence_ids <- pair %>% 
    filter(speaker_id == "Foreign") %>% 
    pull(sentence_id)
  b_sentence_ids <- pair %>% 
    filter(speaker_id == "Native") %>% 
    pull(sentence_id)

@@

```{r eval = FALSE}
## make a directory to store data:
if (!dir.exists("data")) {
  dir.create("data")
}
## If possible, change to OSF URL:
download.file(
  url = "https://github.com/homerhanumat/assignment/raw/master/data/Experiment3.xlsx",
  destfile = "data/Experiment3.xlsx"
)
```

@@

```{r eval = FALSE}
sentences <-
  sentences1 %>% 
  ## pair-id came in as charcter vector, so change to integers:
  mutate(Pair = as.integer(Pair)) %>% 
  ## ditch mst of the columns:
  select(Participant, Pair, Sentence, Accent, Condition, Rating) %>% 
  ## rename most of the variables:
  rename(participant_id = Participant,
         pair_id = Pair,
         sentence_id = Sentence,
         speaker_id = Accent,
         difficulty = Condition,
         rating = Rating) %>% 
  ## change rating scale from [-1, 1] to [100, 100]
  mutate(rating = rating * 100) %>% 
  ## reshape data so that each repreents a single sentence:
  arrange(sentence_id) %>% 
  mutate(participant = rep(1:10, times = 960)) %>% 
  select(-participant_id) %>% 
  spread(key = participant, value = rating, sep = "_rating_")

@@

## Utility function to make matrices of statistics and P-values
## for all pairs of sentences:  one from Speaker A, the other
## from speaker B.
make_diff_matrix <- function(pair, trim) {
  a <- pair %>% 
    filter(speaker_id == "Foreign") %>% 
    select(ends_with("rating")) %>% 
    t() %>% 
    as.matrix()
  b <- pair %>% 
    filter(speaker_id == "Native") %>% 
    select(ends_with("rating")) %>% 
    t() %>% 
    as.matrix()
  mat_statistics <- matrix(0, nrow = ncol(a), ncol = ncol(b))
  mat_pvals <- matrix(0, nrow = ncol(a), ncol = ncol(b))
  for (i in 1:ncol(a)) {
    for (j in 1:ncol(b)) {
      result <- profile_diff(x = a[, i], y = b[, j], trim = trim)
      mat_statistics[i, j] <- result$statistic
      mat_pvals[i, j] <- result$p_value
    }
  }
  rownames(mat_statistics) <- pair %>% 
    filter(speaker_id == "Foreign") %>%
    pull(sentence_id)
  colnames(mat_statistics) <- pair %>% 
    filter(speaker_id == "Native") %>%
    pull(sentence_id)
  rownames(mat_pvals) <- pair %>% 
    filter(speaker_id == "Foreign") %>%
    pull(sentence_id)
  colnames(mat_pvals) <- pair %>% 
    filter(speaker_id == "Native") %>%
    pull(sentence_id)
  list(
    mat_statistics = mat_statistics,
    mat_pvals = mat_pvals
  )
}

@@

means_all <-
  s2 %>% 
  group_by(sentence_id) %>% 
  summarize(mean_rating = mean(rating)) %>% 
  mutate(selected = sentence_id %in% selected_sentences) %>% 
  mutate(selected = ifelse(selected, "used in Corpus", "not used")) %>% 
  inner_join(sentences %>%
               select(sentence_id, speaker_id),
             by = "sentence_id")
```

@@

```{r echo = FALSE, fig.cap = "Boxplots of mean ratings for sentences by foreign and natives speakers."}
means_all %>% 
  ggplot(aes(x = speaker_id, y = mean_rating)) +
  geom_boxplot(fill = "burlywood", outlier.shape = NA) +
  geom_jitter(width = 0.25, size = 0.3) +
  labs(x = "speaker accent", y = "mean rating")
```

@@

```{r echo = FALSE}
means_all %>% 
  group_by(speaker_id) %>% 
  summarize(mean = mean(mean_rating), median = median(mean_rating), n = n()) %>% 
  knitr::kable(caption = "Mean and median of mean ratings for sentences, by type of speaker.")
```

>>>>

make_diff_matrix <- function(pair, trim) {
  a <- pair %>% 
    filter(speaker_accent == "Foreign") %>% 
    select(ends_with("rating")) %>% 
    t() %>% 
    as.matrix()
  b <- pair %>% 
    filter(speaker_accent == "Native") %>% 
    select(ends_with("rating")) %>% 
    t() %>% 
    as.matrix()
  mat_statistics <- matrix(0, nrow = ncol(a), ncol = ncol(b))
  mat_pvals <- matrix(0, nrow = ncol(a), ncol = ncol(b))
  for (i in 1:ncol(a)) {
    for (j in 1:ncol(b)) {
      result <- profile_diff(x = a[, i], y = b[, j], trim = trim)
      mat_statistics[i, j] <- result$statistic
      mat_pvals[i, j] <- result$p_value
    }
  }
  rownames(mat_statistics) <- pair %>% 
    filter(speaker_accent == "Foreign") %>%
    pull(sentence_id)
  colnames(mat_statistics) <- pair %>% 
    filter(speaker_accent == "Native") %>%
    pull(sentence_id)
  rownames(mat_pvals) <- pair %>% 
    filter(speaker_accent == "Foreign") %>%
    pull(sentence_id)
  colnames(mat_pvals) <- pair %>% 
    filter(speaker_accent == "Native") %>%
    pull(sentence_id)
  list(
    mat_statistics = mat_statistics,
    mat_pvals = mat_pvals
  )
}

@@

  ## get sentence ids for each speaker:
  a_sentence_ids <- pair %>% 
    filter(speaker_accent == "Foreign") %>% 
    pull(sentence_id)
  b_sentence_ids <- pair %>% 
    filter(speaker_accent == "Native") %>% 
    pull(sentence_id)

@@

```{r eval = FALSE}
## make a directory to store data:
if (!dir.exists("data")) {
  dir.create("data")
}
## get the excel file:
download.file(
  url = "https://github.com/homerhanumat/assignment/raw/master/data/Experiment3.xlsx",
  destfile = "data/Experiment3.xlsx"
)
```

@@

```{r eval = FALSE}
sentences <-
  sentences1 %>% 
  ## pair-id came in as character vector, so change to integers:
  mutate(Pair = as.integer(Pair)) %>% 
  ## ditch mst of the columns:
  select(Participant, Pair, Sentence, Accent, Condition, Rating) %>% 
  ## rename most of the variables
  rename(participant_id = Participant,
         pair_id = Pair,
         sentence_id = Sentence,
         speaker_accent = Accent,
         difficulty = Condition,
         rating = Rating) %>% 
  ## change rating scale from [-1, 1] to [100, 100]
  mutate(rating = rating * 100) %>% 
  ## reshape data so that each repreents a single sentence:
  arrange(sentence_id) %>% 
  mutate(participant = rep(1:10, times = 960)) %>% 
  select(-participant_id) %>% 
  spread(key = participant, value = rating, sep = "_rating_")

@@

## Utility function
ratings_from_pair <- function(pair) {
  foreign <- pair %>% 
    filter(speaker_accent == "Foreign") %>% 
    select(ends_with("rating")) %>% 
    t() %>% 
    as.matrix()
  native <- pair %>% 
    filter(speaker_accent == "Native") %>% 
    select(ends_with("rating")) %>% 
    t() %>% 
    as.matrix()
  list(foreign = foreign, native = native)
}

## Utility function to make matrices of statistics and P-values
## for all pairs of sentences:  one from Speaker A (Foreign), the other
## from speaker B (Native).
make_diff_matrix <- function(pair, trim) {
  ratings <- ratings_from_pair(pair)
  a <- ratings$foreign
  b <- ratings$native
  mat_statistics <- matrix(0, nrow = ncol(a), ncol = ncol(b))
  mat_pvals <- matrix(0, nrow = ncol(a), ncol = ncol(b))
  for (i in 1:ncol(a)) {
    for (j in 1:ncol(b)) {
      result <- profile_diff(x = a[, i], y = b[, j], trim = trim)
      mat_statistics[i, j] <- result$statistic
      mat_pvals[i, j] <- result$p_value
    }
  }
  rownames(mat_statistics) <- pair %>% 
    filter(speaker_accent == "Foreign") %>%
    pull(sentence_id)
  colnames(mat_statistics) <- pair %>% 
    filter(speaker_accent == "Native") %>%
    pull(sentence_id)
  rownames(mat_pvals) <- pair %>% 
    filter(speaker_accent == "Foreign") %>%
    pull(sentence_id)
  colnames(mat_pvals) <- pair %>% 
    filter(speaker_accent == "Native") %>%
    pull(sentence_id)
  list(
    mat_statistics = mat_statistics,
    mat_pvals = mat_pvals
  )
}

@@

means_all <-
  s2 %>% 
  group_by(sentence_id) %>% 
  summarize(mean_rating = mean(rating)) %>% 
  mutate(selected = sentence_id %in% selected_sentences) %>% 
  mutate(selected = ifelse(selected, "used in Corpus", "not used")) %>% 
  inner_join(sentences %>%
               select(sentence_id, speaker_accent),
             by = "sentence_id")
```

@@

```{r echo = FALSE, fig.cap = "Boxplots of mean ratings for sentences by foreign and natives speakers."}
means_all %>% 
  ggplot(aes(x = speaker_accent, y = mean_rating)) +
  geom_boxplot(fill = "burlywood", outlier.shape = NA) +
  geom_jitter(width = 0.25, size = 0.3) +
  labs(x = "speaker accent", y = "mean rating")
```

@@

```{r echo = FALSE}
means_all %>% 
  group_by(speaker_accent) %>% 
  summarize(mean = mean(mean_rating), median = median(mean_rating), n = n()) %>% 
  knitr::kable(caption = "Mean and median of mean ratings for sentences, by type of speaker.")
```
